Accurate plant pathogen effector protein classification _ab initio_ with **deepredeff**, an ensemble of convolutional neural networks.

### 2020-06-08 main.Rmd
allow tex files in commits

---
title: "Accurate plant pathogen effector protein classification _ab initio_ with **deepredeff**, an ensemble of convolutional neural networks."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm_latex.tex
    toc: no
fontsize: 12pt
fontfamily: sans
papersize: a4paper
author:
  - name: Ruth Kristianingsih, Dan MacLean
  - affiliation: The Sainsbury Laboratory, University of East Anglia, Norwich Research Park, Norwich, UK, NR4 7JJ, dan.maclean@tsl.ac.uk
abstract: "TBD."
geometry: margin=1in
biblio-style: apsr
bibliography: ["r-references.bib", "r-packages.bib"]
graphics: yes
header-includes:
  - \usepackage{amsmath}
  - \usepackage{siunitx}
---

@@

Sequence data were collected from the PHI-Base database version 4.8 [@phibase] by accessing a text dump of the data kindly prepared on request by the PHI-Base team, the file can be accessed at [https://github.com/PHI-base/data/blob/master/releases/phi-base_current.csv](https://github.com/PHI-base/data/blob/master/releases/phi-base_current.csv). The pipeline in Figure \@ref(fig:effectorcollection) outlines the steps used. We filtered plant effector proteins and their taxonomic groups and collected sequences from UniProt Release 2019_05, using the code in [https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-data-new/binary-class/0001_first_step_getting_data.Rmd](https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-data-new/binary-class/0001_first_step_getting_data.Rmd). We created a correspondingly sized data set of non-effectors with secretion signals originating in species matched to those from which the effectors were drawn. We downloaded sequences for randomly selected proteins matching these criteria from Ensembl databases [@yates2020] : specifically Ensembl Fungi, Protists and Bacteria manually using the BioMart tools [@smedleybiomart2009]. Since the BioMart tool is not available on Ensembl Bacteria, we downloaded whole proteome protein sequnces from species matched to those from which the effector came using FTP. With these we used SignalP 3.0 [@dyrlovbendtsen2004] in order to filter the secreted sequences and selected accordingly. We used default paramaters from SignalP, except the type of organism group which is `euk` for both fungi and oomycete sequences, and `gram-` or `gram+` for bacteria sequences. Redundant sequences were filtered using BLASTp [@camachoblast2009]. We achieved these steps using the code in [https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-secreted-data/0005_process_signalp_data.Rmd](https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-secreted-data/0005_process_signalp_data.Rmd).

```{r effectorcollection, out.width='50%', fig.align='center', fig.cap="Workflow diagram for collection of effector sequences from the PHI-Base database annotation and cross reference to UniProt"}
knitr::include_graphics(
  "figures/flowchart_effectorsequence_collection.pdf"
)
```

@@

@@

```{r, bacefftable, echo=FALSE, message = FALSE}

@@

```{r hyperparam}
 readr::read_tsv("tables/hyperparamscan_info.tab") %>%
  knitr::kable( "latex", longtable = T, booktabs = T,
           digits = 0,
  align = 'c',
  caption = 'Hyperparameters used in automatic scanning of hyperparameter space and best performing parameters values for each model'
         ) %>%
  kableExtra::kable_styling(latex_options = c("repeat_header"), font_size = 7)

>>>>

---
title: "Accurate plant pathogen effector protein classification _ab initio_ with **deepredeff**, an ensemble of convolutional neural networks."
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  bookdown::pdf_document2:
    citation_package: natbib
    keep_tex: true
    fig_caption: true
    latex_engine: pdflatex
    template: svm_latex.tex
    toc: no
    includes:
      in_header: preamble.tex
fontsize: 12pt
fontfamily: sans
papersize: a4paper
author:
  - name: Ruth Kristianingsih, Dan MacLean
  - affiliation: The Sainsbury Laboratory, University of East Anglia, Norwich Research Park, Norwich, UK, NR4 7JJ, dan.maclean@tsl.ac.uk
abstract: "TBD."
geometry: margin=1in
biblio-style: apsr
bibliography: ["r-references.bib", "r-packages.bib"]
graphics: yes
header-includes:
  - \usepackage{amsmath}
  - \usepackage{siunitx}
---

@@

Sequence data were collected from the PHI-Base database version 4.8 [@phibase] by accessing a text dump of the data prepared on request by the PHI-Base team, the file can be accessed at [https://github.com/PHI-base/data/blob/master/releases/phi-base_current.csv](https://github.com/PHI-base/data/blob/master/releases/phi-base_current.csv). The pipeline in Figure \@ref(fig:effectorcollection) outlines the steps used. We filtered plant effector proteins and their taxonomic groups and collected sequences from UniProt Release 2019_05, using the code in [https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-data-new/binary-class/0001_first_step_getting_data.Rmd](https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-data-new/binary-class/0001_first_step_getting_data.Rmd). We created a correspondingly sized data set of non-effectors with secretion signals originating in species matched to those from which the effectors were drawn. We downloaded sequences for randomly selected proteins matching these criteria from Ensembl databases [@yates2020] : specifically Ensembl Fungi, Protists and Bacteria manually using the BioMart tools [@smedleybiomart2009]. Since the BioMart tool is not available on Ensembl Bacteria, we downloaded whole proteome protein sequnces from species matched to those from which the effector came using FTP. With these we used SignalP 3.0 [@dyrlovbendtsen2004] in order to filter the secreted sequences and selected accordingly. We used default paramaters from SignalP, except the type of organism group which is `euk` for both fungi and oomycete sequences, and `gram-` or `gram+` for bacteria sequences. Redundant sequences were filtered using BLASTp [@camachoblast2009]. We achieved these steps using the code in [https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-secreted-data/0005_process_signalp_data.Rmd](https://github.com/TeamMacLean/ruth-effectors-prediction/blob/master/scripts/r-scripts/getting-secreted-data/0005_process_signalp_data.Rmd).

@@

-->
```{r effectorcollection, out.width='50%', fig.align='center', fig.cap="Workflow diagram for collection of effector sequences from the PHI-Base database annotation and cross reference to UniProt"}
knitr::include_graphics(
  "figures/flowchart_effectorsequence_collection.pdf"
)
```

@@

\beginsupplement
```{r, bacefftable, echo=FALSE, message = FALSE}

@@

```{r, hyperparam}
 readr::read_tsv("tables/hyperparamscan_info.tab") %>%
  knitr::kable( "latex", longtable = T, booktabs = T,
           digits = 0,
  align = 'c',
  caption = 'Hyperparameters used in automatic scanning of hyperparameter space and best performing parameters values for each model'
         ) %>%
  kableExtra::kable_styling(latex_options = c("repeat_header"), font_size = 7)
